{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2025-09-23 poincare.tex
I wonder if the expression in exercise 3.6 might be wrong. I have added a proposed solution to the exercise, hoping it might be corrected if I am wrong.

\begin{Exercise}[]
Use Eq.~(\ref{eq:exp_map_Lorentz}) to write out an explicit expression for a Lorentz boost in the $x$-direction with rapidity $\eta$.
\end{Exercise}
\begin{Answer} 
For a Lorentz boost in the $x$-direction all parameters are zero except $\omega_{10}=-\omega_{01}=\eta$. This gives
\begin{equation*}
\Lambda = \exp(i\omega^{10}M_{10})= \exp(i\eta K_1)=\cosh(i\eta K_1)+\sinh(i\eta K_1).
\end{equation*}
where we have used the relationship between the formal power series for the exponential and hyperbolic functions. Now the hyperbolic cosine holds the terms even in $K_1$ and hyperbolic sine the odd terms. From
\[iK_1 = - \left[\begin{matrix} 0 & 1  & 0 &  0\\  1 & 0 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right], \quad
(iK_1)^2 = \left[\begin{matrix} 1 & 0  & 0 &  0\\  0 & 1 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right], \quad
(iK_1)^3 = -\left[\begin{matrix} 0 & 1  & 0 &  0\\  1 & 0 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right], \]
we can write
\begin{equation*}
\Lambda = I+ (\cosh\eta-1)\left[\begin{matrix} 1 & 0  & 0 &  0\\  0 & 1 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right]-\sinh\eta\left[\begin{matrix} 0 & 1  & 0 &  0\\  1 & 0 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right]
=\left[\begin{matrix} \cosh\eta & -\sinh\eta  & 0 & 0 \\  -\sinh\eta & \cosh\eta & 0 & 0 \\  0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{matrix}\right].
\end{equation*}
Thus we are left with the standard matrix for Lorentz transformations in the $x$-direction (using rapidity).
\end{Answer}

\begin{Exercise}[]
Show that a general boost in the direction of the unit vector $\mathbf n$ with rapidity $\eta$ can be written as
\begin{equation*}
B(\eta,\mathbf n)=I+(\cosh\eta-1)(\mathbf n \cdot \mathbf K)^2-i\sinh\eta (\mathbf n \cdot \mathbf K).
\end{equation*}
\end{Exercise}

>>>>

\begin{Exercise}[label=ex:Lorentz_explicit_x]
Use Eq.~(\ref{eq:exp_map_Lorentz}) to write out an explicit expression for a Lorentz boost in the $x$-direction with rapidity $\eta$.
\end{Exercise}
\begin{Answer} 
For a Lorentz boost in the $x$-direction all parameters are zero except $\omega_{10}=-\omega_{01}=\eta$. This gives
\begin{equation*}
\Lambda = \exp(i\omega^{10}M_{10})= \exp(i\eta K_1)=\cosh(i\eta K_1)+\sinh(i\eta K_1).
\end{equation*}
where we have used the relationship between the formal power series for the exponential and hyperbolic functions. Now the hyperbolic cosine holds the terms even in $K_1$ and hyperbolic sine the odd terms. From
\[iK_1 = - \left[\begin{matrix} 0 & 1  & 0 &  0\\  1 & 0 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right], \quad
(iK_1)^2 = \left[\begin{matrix} 1 & 0  & 0 &  0\\  0 & 1 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right], \quad
(iK_1)^3 = -\left[\begin{matrix} 0 & 1  & 0 &  0\\  1 & 0 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right], \]
we can write
\begin{equation*}
\Lambda = I+ (\cosh\eta-1)\left[\begin{matrix} 1 & 0  & 0 &  0\\  0 & 1 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right]-\sinh\eta\left[\begin{matrix} 0 & 1  & 0 &  0\\  1 & 0 & 0 & 0 \\  0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \end{matrix}\right]
=\left[\begin{matrix} \cosh\eta & -\sinh\eta  & 0 & 0 \\  -\sinh\eta & \cosh\eta & 0 & 0 \\  0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{matrix}\right].
\end{equation*}
Thus we are left with the standard matrix for Lorentz transformations in the $x$-direction (using rapidity).
\end{Answer}

\begin{Exercise}[]
Show that a general boost in the direction of the unit vector $\mathbf n$ with rapidity $\eta$ can be written as
\begin{equation*}
B(\eta,\mathbf n)=I+(\cosh\eta-1)(\mathbf n \cdot \mathbf K)^2-i\sinh\eta (\mathbf n \cdot \mathbf K).
\end{equation*}
\end{Exercise}
\begin{Answer}
The boost generators are given by
\begin{align*}
    iK_1=- \begin{bmatrix}
        0 & 1 & 0 & 0\\
        1 & 0 & 0 & 0\\
        0 & 0 & 0 & 0\\
        0 & 0 & 0 & 0
    \end{bmatrix}, \quad
    iK_2=- \begin{bmatrix}
        0 & 0 & 1 & 0\\
        0 & 0 & 0 & 0\\
        1 & 0 & 0 & 0\\
        0 & 0 & 0 & 0
    \end{bmatrix}, \quad
    iK_3=- \begin{bmatrix}
        0 & 0 & 0 & 1\\
        0 & 0 & 0 & 0\\
        0 & 0 & 0 & 0\\
        1 & 0 & 0 & 0
    \end{bmatrix},
\end{align*}
thus we have (using $n_x^2+n_y^2+n_z^2=1$):
\begin{align*}
    i \vec{n}\cdot\vec{K} &= -\begin{bmatrix}
        0 & n_x & n_y & n_z \\
        n_x & 0 & 0 & 0 \\
        n_y & 0 & 0 & 0 \\
        n_z & 0 & 0 & 0
    \end{bmatrix},
    \\
    (i \vec{n} \cdot \vec{K})^2 &= \begin{bmatrix}
        1 & 0 & 0 & 0\\
        0 & n_x^2 & n_xn_y & n_xn_z \\
        0 & n_xn_y & n_y^2 & n_yn_z \\
        0 & n_xn_z & n_yn_z & n_z^2
    \end{bmatrix}
    \\
    (i \vec{n} \cdot \vec{K})^3 &= - \begin{bmatrix}
        0 & n_x & n_y & n_z \\
        n_x & 0 & 0 & 0 \\
        n_y & 0 & 0 & 0 \\
        n_z & 0 & 0 & 0
    \end{bmatrix} = \vec{n}\cdot\vec{K}.
\end{align*}
We see that $(i \vec{n}\cdot\vec{K})^{2k+1}= i \vec{n}\cdot\vec{K}$ and $(i \vec{n}\cdot\vec{K})^{2k}=(i \vec{n}\cdot\vec{K})^2$. Using this we find (setting $\omega^{00}=\omega^{ij}=0$ and $\omega^{i0}=-\omega^{0i}=n^i \eta$)
\begin{align*}
    B(\eta,\vec{n}) &= \exp\Bigl(\frac{i}{2}\omega^{\rho\sigma}M_{\rho\sigma}\Bigr)
    \\
    &= \exp \bigl(\eta (i \vec{n}\cdot\vec{K})\bigr)
    \\
    &= I+\eta(i\vec{n}\cdot\vec{K})+\frac{1}{2}\eta^2(i\vec{n}\cdot\vec{K})^2+\cdots
    \\
    &= I+\frac{1}{2}\eta^2(i \vec{n}\cdot\vec{K})^2+\frac{1}{4!}\eta^4(i\vec{n}\cdot\vec{K})^4+\cdots
    \\ &\qquad +\eta(i\vec{n}\cdot\vec{K})+\frac{1}{3!}\eta^3(i\vec{n}\cdot\vec{K})^3+\cdots
    \\
    &= I + \Bigl(\frac{1}{2}\eta^2+\frac{1}{4!}\eta^4+\cdots\Bigr)(i\vec{n}\cdot\vec{K})^2+\Bigl(\eta+\frac{1}{3!}\eta^3+\cdots\Bigr)(i\vec{n}\cdot\vec{K})
    \\
    &= I + (\cosh\eta-1)(i\vec{n}\cdot\vec{K})^2+\sinh\eta (i\vec{n}\cdot\vec{K})
    \\
    &= I - (\cosh\eta-1)(\vec{n}\cdot\vec{K})^2+i\sinh\eta(\vec{n}\cdot\vec{K}),
\end{align*}
which is \emph{almost} the expression given in the text. Setting $\vec{n}=\vec{e_x}$ should give the result from Ex.~\ref{ex:Lorentz_explicit_x}:
\begin{align*}
    B(\eta,\vec{e_x})&=I+(\cosh\eta-1)(iK_1)^2+\sinh\eta(iK_1)
    \\
    &= I+(\cosh\eta-1)\begin{bmatrix}
        1&0&0&0 \\ 0&1&0&0 \\ 0&0&0&0 \\ 0&0&0&0
    \end{bmatrix} -\sinh\eta \begin{bmatrix}
        0&1&0&0 \\ 1&0&0&0 \\ 0&0&0&0 \\ 0&0&0&0
    \end{bmatrix},
\end{align*}
which is the answer to Ex.~\ref{ex:Lorentz_explicit_x}. Since these results coincide, I suspect there is an error in this exercise, and that the answer really should be
\begin{equation*}
    B(\eta,\vec{n}) = I + (\cosh\eta-1)(i\vec{n}\cdot\vec{K})^2+\sinh\eta (i\vec{n}\cdot\vec{K}).
\end{equation*}
For comparison, the expression given in the problem gives
\begin{align*}
    B(\eta,\vec{e_x}) &= I+(\cosh\eta-1) K_1^2-i\sinh(\eta) K_1
    \\
    &= I - (\cosh\eta-1)\begin{bmatrix}
        1&0&0&0 \\ 0&1&0&0 \\ 0&0&0&0 \\ 0&0&0&0
    \end{bmatrix} +\sinh\eta \begin{bmatrix}
        0&1&0&0 \\ 1&0&0&0 \\ 0&0&0&0 \\ 0&0&0&0
    \end{bmatrix}.
\end{align*}
\end{Answer}
